Endogenous long-run output impacts of monetary and fiscal policy: a growth theory approach

### 2021-12-04 paper.tex
added actions and more work on paper

Here \(\gamma_0\) is the non-intervention (baseline) rate of growth in research labor, \(\gamma_1\) is the sensitivity of labor to the prevailing interest rate's deviation from the ``natural'' rate \(i^*\), and \(\gamma_2\) is the proportion of government spending that is dedicated to research and development.  For instance, federal grants to universities would be considered spending on basic research.   
\\
\\
\(L_1\) is determined by the above equations:
\[\frac{dL_1(t)}{dt}=\frac{dL}{dt}-\frac{dL_2(t)}{dt}-\frac{dL_3(t)}{dt}\] 
\[=\eta\left(L_1(t)+L_2(t)+L_3(t)\right)-\frac{dL_2(t)}{dt}-\frac{dL_3(t)}{dt}\]
\\
\\
Technology dynamics are directly proportional to the labor dediated to research.  While in reality research requires some level of capital, the assumption simplifies the analysis.  Many kinds of research (for example, mathematics and economics) requires minimal capital investment.  Of course, the physical sciences often require massive capital investments.  This is a shortcoming of the model.
\[\frac{dA(t)}{dt}=\beta L_2(t)\]

@@

\begin{figure}[htb]
\begin{framed}
Densities of the loss distribution of \(X_T\) with \(10,000\) assets.  Each asset has \(p=.03\) and exposure drawn from a Gamma\((10, 1000)\) distribution to ensure heterogeneous exposures. The densities are normalized to be a percentage of total exposure. \newline
\begin{minipage}[t]{.48\textwidth}
\centering
\includegraphics[width=1\textwidth]{output_per_labor_100}
\caption{\(\alpha=.3\), \(\sigma=.5\), \(\lambda=0\), \(q=0\).  \label{fig5}}
\end{minipage}\hfill
\begin{minipage}[t]{.48\textwidth}
\centering
\includegraphics[width=1\textwidth]{output_per_labor_100}
\caption{\(z_0=1\), \(T=1\)  \label{fig6}}
\end{minipage}\hfill
\end{framed}
\end{figure}

\section{Numerical Inversion}

>>>>

Here \(\gamma_0\) is the non-intervention (baseline) rate of growth in research labor, \(\gamma_1\) is the sensitivity of labor to the prevailing interest rate's deviation from the ``natural'' rate \(i^*\), and \(\gamma_2\) is the proportion of government spending that is dedicated to research and development.  For instance, federal grants to universities would be considered spending on basic research.   
\\
\\
\(L_1\) is determined by the above equations:
\[\frac{dL_1(t)}{dt}=\frac{dL}{dt}-\frac{dL_2(t)}{dt}-\frac{dL_3(t)}{dt}\] 
\[=\eta\left(L_1(t)+L_2(t)+L_3(t)\right)-\frac{dL_2(t)}{dt}-\frac{dL_3(t)}{dt}\]
\\
\\
Technology dynamics are directly proportional to the labor dediated to research.  While in reality research requires some level of capital, the assumption simplifies the analysis.  Many kinds of research (for example, mathematics and economics) requires minimal capital investment.  Of course, the physical sciences often require massive capital investments.  This is a shortcoming of the model.  Note that this also deviates from the classic Solow model where technological growth is exogoneous.
\[\frac{dA(t)}{dt}=\beta L_2(t)\]

@@

\subsection{Model analysis}

I focus on output per capita, \(\frac{Y(t)}{L(t)}\). There are three possible long-run states to the model.  First is exponential growth.  This occurs if \(A(t)\) is allowed to grow through time while not crowding out growth in \(L_1(t)\).  The second is a steady state.  This occurs if \(L_2(t)\) shrinks to zero, causing \(A(t)\) to converge to a constant.  The third is economic collapse.  This occurs if \(L_1(t)\) shrinks to zero due to crowding out from \(L_2(t)\) and \(L_3(t)\).

\subsubsection{Exponential growth}
If \(\gamma_0+\gamma_1 (i^*-i)+\gamma_2 g > 0\) then \(L_2\) will continue to grow.  This implies that \(A(t)\) will grow without bound, and continued technological growth can continue to provide improved output per capita.  However, this is only the case if both \(L_2(t)\) grows slowly enough to not cause capital to be consumed and \(L_1(t)\) also continues to grow.

\subsection{Steady state}
If \(\gamma_0+\gamma_1 (i^*-i)+\gamma_2 g <0\) then \(L_2(t)\) will shrink to zero and \(A(t)\) will converge to a constant.  Much like the original Solow model, the economy reaches a long-run steady state.

\subsection{Economic collapse}
If either \(L_2\) or \(L_3\) grow too large, the \(K\) can shrink to zero or \(L_1\) can shrink to zero.  This can occur if interest rates are held too low or \(g\) is too large.

\section{Results}

Since the model is a system of non-linear differential equations, the model must be solved numerically.

\begin{figure}[htb]
\begin{framed}

\begin{minipage}[t]{.48\textwidth}
\centering
\includegraphics[width=1\textwidth]{economy1_100}
\end{minipage}\hfill
\begin{minipage}[t]{.48\textwidth}
\centering
\includegraphics[width=1\textwidth]{economy2_100}
\end{minipage}\hfill
\end{framed}
\end{figure}
